### 2014-11-19 doc/thesis/evaluation.tex
Updating the evaluation data to include the inter category
Adding raw text about related work

\begin{table}[!ht]
\begin{center}
\begin{tabular}{|l|c|c|c|c|c|c|c|}
\hline
\textbf{Case study} & \textbf{easy} & \textbf{poly} & \textbf{inter} & \textbf{hard} & \textbf{Total} & \textbf{\%} \\
\hline
Lua Standard Library & 87 & 6 & 0 & 36 & 129 & 67\% \\
\hline
MD5 & 13 & 0 & 0 & 0 & 13 & 100\% \\
\hline
LuaSocket & 109 & 1 & 0 & 13 & 123 & 89\% \\
\hline
HTTP Digest & 0 & 0 & 0 & 1 & 1 & 0\% \\
\hline
Typical & 1 & 0 & 0 & 0 & 1 & 100\% \\
\hline
Modulo 11 & 7 & 0 & 0 & 2 & 9 & 78\% \\
\hline
Typed Lua Compiler & 144 & 0 & 0 & 10 & 154 & 94\% \\
\hline
\end{tabular}
\end{center}
\caption{Evaluation results for each case study}
\label{tab:evalbycase}
\end{table}

@@

\begin{table}[!ht]
\begin{center}
\begin{tabular}{|c|c|c|c|c|c|c|}
\hline
\textbf{Module} & \textbf{easy} & \textbf{poly} & \textbf{inter} & \textbf{hard} & \textbf{Total} & \textbf{\%} \\
\hline
base & 9 & 1 & 0 & 16 & 26 & 35\% \\
\hline
coroutine & 1 & 0 & 0 & 6 & 7 & 14\% \\
\hline
package & 5 & 0 & 0 & 3 & 8 & 62\% \\
\hline
string & 14 & 0 & 0 & 2 & 16 & 88\% \\
\hline
utf8 & 6 & 0 & 0 & 0 & 6 & 100\% \\
\hline
table & 1 & 5 & 0 & 1 & 7 & 14\% \\
\hline
math & 26 & 0 & 0 & 1 & 27 & 96\% \\
\hline
io & 15 & 0 & 0 & 6 & 21 & 71\% \\
\hline
os & 10 & 0 & 0 & 1 & 11 & 91\% \\
\hline
\end{tabular}
\end{center}
\caption{Evaluation results for Lua Standard Library}
\label{tab:evallsl}
\end{table}

@@

\begin{table}[!ht]
\begin{center}
\begin{tabular}{|c|c|c|c|c|c|c|}
\hline
\textbf{Module} & \textbf{easy} & \textbf{poly} & \textbf{inter} & \textbf{hard} & \textbf{Total} & \textbf{\%} \\
\hline
socket & 50 & 0 & 0 & 10 & 60 & 83\% \\
\hline
ftp & 5 & 0 & 0 & 1 & 6 & 83\% \\
\hline
http & 4 & 0 & 0 & 1 & 5 & 80\% \\
\hline
smtp & 6 & 0 & 0 & 1 & 7 & 86\% \\
\hline
mime & 17 & 0 & 0 & 0 & 17 & 100\% \\
\hline
ltn12 & 19 & 1 & 0 & 0 & 20 & 95\% \\
\hline
url & 8 & 0 & 0 & 0 & 8 & 100\% \\
\hline
\end{tabular}
\end{center}
\caption{Evaluation results for LuaSocket}
\label{tab:evalsocket}
\end{table}

@@

\begin{table}[!ht]
\begin{center}
\begin{tabular}{|c|c|c|c|c|c|c|}
\hline
\textbf{Module} & \textbf{easy} & \textbf{poly} & \textbf{inter} & \textbf{hard} & \textbf{Total} & \textbf{\%} \\
\hline
http-digest & 0 & 0 & 0 & 1 & 1 & 0\% \\
\hline
\end{tabular}
\end{center}
\caption{Evaluation results for HTTP Digest}
\label{tab:evalhttpdigest}
\end{table}

@@

\begin{table}[!ht]
\begin{center}
\begin{tabular}{|c|c|c|c|c|c|c|}
\hline
\textbf{Module} & \textbf{easy} & \textbf{poly} & \textbf{inter} & \textbf{hard} & \textbf{Total} & \textbf{\%} \\
\hline
tlast & 46 & 0 & 0 & 1 & 47 & 98\% \\
\hline
tltype & 65 & 0 & 0 & 0 & 65 & 100\% \\
\hline
tlst & 26 & 0 & 0 & 0 & 26 & 100\% \\
\hline
tllexer & 2 & 0 & 0 & 9 & 11 & 18\% \\
\hline
tlparser & 1 & 0 & 0 & 0 & 1 & 100\% \\
\hline
tldparser & 1 & 0 & 0 & 0 & 1 & 100\% \\
\hline
tlchecker & 2 & 0 & 0 & 0 & 2 & 100\% \\
\hline
tlcode & 1 & 0 & 0 & 0 & 1 & 100\% \\
\hline
\end{tabular}
\end{center}
\caption{Evaluation results for Typed Lua Compiler}
\label{tab:evaltlc}
\end{table}

@@

\begin{itemize}
\item \citep{tidallock}
\item \citep{bonnaire-sergeant2012typed-clojure}
\item \citep{vitousek2014deg}
\item \citep{allende2013gts}
\item \citep{tobin-hochstadt2008ts} 
\item \citep{dart}
\item \citep{typescript}
\item \citep{bierman2014typescript}
\item \citep{politz2012semantics}
\end{itemize}

>>>>

\begin{table}[!ht]
\begin{center}
\begin{tabular}{|l|c|c|c|c|c|c|c|}
\hline
\textbf{Case study} & \textbf{easy} & \textbf{poly} & \textbf{inter} & \textbf{hard} & \textbf{Total} & \textbf{\%} \\
\hline
Lua Standard Library & 80 & 6 & 9 & 34 & 129 & 62\% \\
\hline
MD5 & 13 & 0 & 0 & 0 & 13 & 100\% \\
\hline
LuaSocket & 109 & 1 & 2 & 11 & 123 & 89\% \\
\hline
HTTP Digest & 0 & 0 & 1 & 0 & 1 & 0\% \\
\hline
Typical & 1 & 0 & 0 & 0 & 1 & 100\% \\
\hline
Modulo 11 & 7 & 0 & 0 & 2 & 9 & 78\% \\
\hline
Typed Lua Compiler & 144 & 0 & 1 & 9 & 154 & 94\% \\
\hline
\end{tabular}
\end{center}
\caption{Evaluation results for each case study}
\label{tab:evalbycase}
\end{table}

@@

\begin{table}[!ht]
\begin{center}
\begin{tabular}{|c|c|c|c|c|c|c|}
\hline
\textbf{Module} & \textbf{easy} & \textbf{poly} & \textbf{inter} & \textbf{hard} & \textbf{Total} & \textbf{\%} \\
\hline
base & 9 & 1 & 3 & 13 & 26 & 35\% \\
\hline
coroutine & 1 & 0 & 0 & 6 & 7 & 14\% \\
\hline
package & 5 & 0 & 0 & 3 & 8 & 62\% \\
\hline
string & 12 & 0 & 0 & 4 & 16 & 75\% \\
\hline
utf8 & 6 & 0 & 0 & 0 & 6 & 100\% \\
\hline
table & 1 & 5 & 0 & 1 & 7 & 14\% \\
\hline
math & 22 & 0 & 4 & 1 & 27 & 81\% \\
\hline
io & 15 & 0 & 0 & 6 & 21 & 71\% \\
\hline
os & 9 & 0 & 2 & 0 & 11 & 82\% \\
\hline
\end{tabular}
\end{center}
\caption{Evaluation results for Lua Standard Library}
\label{tab:evallsl}
\end{table}

@@

\begin{table}[!ht]
\begin{center}
\begin{tabular}{|c|c|c|c|c|c|c|}
\hline
\textbf{Module} & \textbf{easy} & \textbf{poly} & \textbf{inter} & \textbf{hard} & \textbf{Total} & \textbf{\%} \\
\hline
socket & 50 & 0 & 0 & 10 & 60 & 83\% \\
\hline
ftp & 5 & 0 & 1 & 0 & 6 & 83\% \\
\hline
http & 4 & 0 & 1 & 0 & 5 & 80\% \\
\hline
smtp & 6 & 0 & 0 & 1 & 7 & 86\% \\
\hline
mime & 17 & 0 & 0 & 0 & 17 & 100\% \\
\hline
ltn12 & 19 & 1 & 0 & 0 & 20 & 95\% \\
\hline
url & 8 & 0 & 0 & 0 & 8 & 100\% \\
\hline
\end{tabular}
\end{center}
\caption{Evaluation results for LuaSocket}
\label{tab:evalsocket}
\end{table}

@@

\begin{table}[!ht]
\begin{center}
\begin{tabular}{|c|c|c|c|c|c|c|}
\hline
\textbf{Module} & \textbf{easy} & \textbf{poly} & \textbf{inter} & \textbf{hard} & \textbf{Total} & \textbf{\%} \\
\hline
http-digest & 0 & 0 & 1 & 0 & 1 & 0\% \\
\hline
\end{tabular}
\end{center}
\caption{Evaluation results for HTTP Digest}
\label{tab:evalhttpdigest}
\end{table}

@@

\begin{table}[!ht]
\begin{center}
\begin{tabular}{|c|c|c|c|c|c|c|}
\hline
\textbf{Module} & \textbf{easy} & \textbf{poly} & \textbf{inter} & \textbf{hard} & \textbf{Total} & \textbf{\%} \\
\hline
tlast & 46 & 0 & 1 & 0 & 47 & 98\% \\
\hline
tltype & 65 & 0 & 0 & 0 & 65 & 100\% \\
\hline
tlst & 26 & 0 & 0 & 0 & 26 & 100\% \\
\hline
tllexer & 2 & 0 & 0 & 9 & 11 & 18\% \\
\hline
tlparser & 1 & 0 & 0 & 0 & 1 & 100\% \\
\hline
tldparser & 1 & 0 & 0 & 0 & 1 & 100\% \\
\hline
tlchecker & 2 & 0 & 0 & 0 & 2 & 100\% \\
\hline
tlcode & 1 & 0 & 0 & 0 & 1 & 100\% \\
\hline
\end{tabular}
\end{center}
\caption{Evaluation results for Typed Lua Compiler}
\label{tab:evaltlc}
\end{table}

@@

Tidal Lock \citep{tidallock} is a prototype of an optional type system for Lua.
It also includes the refinement of table types.
In fact, Tidal Lock inspired Typed Lua to also include this feature.
However, Tidal Lock covers a little subset of Lua.
Statements include declaration of local variables, multiple assignments,
function application, and the return statement, that is, they do not
include any control-flow statements.
Expressions include primitive literals, table indexing, function application,
function declaration, and the table constructor, but they do not include
binary operations.

The table type from Tidal Lock can only represent records, that is,
it cannot describe hash tables and arrays yet, though we can refine them.
Tidal Lock also includes field types to describe the type of the fields
of a table type.
The field types describe whether a table field is mutable or imutable
in a table type.
Field types are the feature that allow the refinement of table types in
Tidal Lock.

Tidal Lock is similar to Typed Lua because it is also a structural type
system based on subtyping.
It also uses local type inference to deduce the type of unannotated expressions,
and have a dynamic type.
However, Tidal Lock does not support union types, recursive types, and variadic types.
It also does not support any object-oriented idiom.

Typed Racket \citep{tobin-hochstadt2008ts} is a statically typed version
of the Racket language, which is a Scheme dialect.
The main purpose of Typed Racket is to allow programmers to combine
untyped modules, which are written in Racket, with typed modules, which are
written in Typed Racket.
It also uses local type inference to deduce the type of unannotated expressions.

The type system of Typed Racket introduced \emph{occurrence typing}
\citep{tobin-hochstadt2010ltu}.
It is a novel way to use type predicates in control flow statements
to refine union types.
The type system is rich and includes basic types, function types,
union types, recursive types, and structure types.
Function types can return multiple values.
Structure types are similar to our interfaces, as they describe
record types.
The type system is also structural and based on subtyping.
It also includes the dynamic type \texttt{Any}, which is the top
type in the system.
Typed Racket also supports polymorphic functions and data structures.

Typed Clojure \citep{bonnaire-sergeant2012typed-clojure} is an
optional type system for Clojure.
Although Clojure is a Lisp dialect that runs on the Java Virtual Machine,
Common Language Runtime, and JavaScript, Typed Clojure runs only on
the Java Virtual Machine.
Perhaps, this restriction pushed Typed Clojure to support Java classes
and some Java types such as \texttt{Long}, \texttt{Double}, and \texttt{String}.
Typed Clojure also provides optional type annotations and uses
local type inference to deduce the type of unannotated expressions.
It also assigns the type \texttt{Any} to unannotated function parameters,
which is the top type in the type system.

The type system of Typed Clojure is rich.
It includes polymorphic function types, union types, intersection types,
lists, vectors, maps, sets, and recursive types.
Function types can also have rest parameters, which are similar
to our variadic types, but can only appear on the input parameter
of function types.
In fact, its function types cannot return multiple results.
It also uses ocurrence typing to allow control flow statements to
refine union types.
The type system is also structural and based on subtyping.

Dart \citep{dart} is a new class-based object-oriented programming
language.
It includes optional type annotations and compiles to JavaScript.
The type system of Dart is nomial and includes base types,
function types, lists, and maps.
It also supports generics, and the programmer can define
generic functions, lists, and maps.

Even though Dart has optional typing and static types by
default do not affect run-time semantics, it has an
execution mode that affects run-time.
The \emph{checked mode} inserts run-time assertions that
verifies whether static types match run-time tags.
The \emph{production mode} is the default execution mode
that does not include any assertions.

TypeScript \citep{typescript} is a JavaScript extension
that includes optional type annotations and class-based
object-oriented programming.
It also uses local type inference to deduce the type
of unannotated expressions.
The type system of TypeScript is structural and based
on subtyping.
It includes the dynamic type, primitive types, function
types, array types, tuple types, and recursive types.
TypeScript also support generics.

Even though TypeScript is unsound by design,
\citet{bierman2014typescript} shows how to make TypeScript sound.
They use a reduced core of TypeScript to formalize a
sound type system for TypeScript, but also to formalize
its current unsound type system.

\citet{politz2012semantics} proposes semantics and types for objects
with first-class member names, a well-known feature from scripting languages.
Their type system uses string patterns to describe the members of
an object, and define a complex subtyping relation to validate
these patterns.
They also provide an implementation of their system to JavaScript.

Gradualtalk \citep{allende2013gts} is a Smalltalk dialect that
supports gradual typing.
The type system combines nominal and structural typing.
It includes function types, union types, structural types,
nominal types, a self type, and parametric polymorphism.
The type system also relies on subtyping and consistent-subtyping.

Gradualtalk inserts run-time checks that ensure dynamically
typed code does not violate statically typed code.
\citet{allende2013cis} perform a carefull evaluation about
cast insertion on Gradualtalk.
They report that usually cast insertions impact on execution
performance.

Reticulated Python \citep{vitousek2014deg} is a Python compiler
that supports gradual typing.
The type system is structural and based on subtyping.
It includes base types, the dynamic type, list types,
dictionary types, tuple types, function types, set types,
object types, class types, and recursive types.
It includes class and object types to differentiate the
type of class declarations and instances.
It also uses local type inference.
Besides static type checking, Reticulated Python also introduces
three different approaches for inserting run-time assertions.

Hack \citep{hack} is a new programming language that combines
static and dynamic type in PHP code.
Its type system includes generics, nullable types, collections,
and function types.

The Ruby Type Checker \citep{ren2013rtc} is a library that
performs type checking during run-time.
The library provides type annotations that the programmer
can use on classes and methods.
Its type system includes nominal types, union types,
intersection types, method types, parametric polymorphism,
and type casts.
